---
title: ""
output: html_document
---

&nbsp;

<center>

# Correlations

</center>

&nbsp;

----

&nbsp;

## Purpose 
Enhance your analysis further with

- 1.0 Pearson Correlations
- 2.0 Lag Correlations
- 3.0 Correlation Scatterplots (all in one instead of individually generated)
- 4.0 Correlation Funnels


```{r setup, include=FALSE}
# set working directory to project folder
root.dir = rprojroot::find_rstudio_root_file()
setwd(root.dir)

# *************************************************
# 1.0 Imports ----
# *************************************************
# Input libraries
library(readr) #read csv
library(readxl) # for retrieving and loading data

# Standard libraries
library(purrr) # for map_df
library(dplyr) # for glimpse
library(psych) # for describe
library(tidyr) # for gather
library(stepR) # for formatting data
library(moments) # for analyzing skewness
library(anomalize) # for anomaly detection

# Transformation libraries
library(caret) # for standardization, preprocess
library(recipes) # for formatting data, creating new variables, onehotencode
library(lubridate) # for changing char column to date data type

# Modelling libraries
library(timetk) #time series
library(tidymodels)
library(factoextra)
library(performance)
library(leaps)
library(forcats)
library(tidyquant)
library(rpart)
library(rpart.plot)

# Visualization libraries and theme
library(ggstatsplot) #for correlation matrix
library(GGally) #for data understanding, scatterplot
library(ggplot2) #for plots
library(ggpubr) #stitching plots
theme_set(theme_pubr())

# *************************************************
# 2.0 Input & Outputs ----
# *************************************************
# clear all variables in work space
rm(list = ls())

# input
sample_data_filename = "/Users/shellys/github/cm-mks-ds-starters-correlation/00_data/website_performance_daily.csv"

# *************************************************
# 3.0 Load Data ----
# *************************************************
sample_data_tbl = read_csv(sample_data_filename)

```

&nbsp;

----

&nbsp;

## 1.0 Pearson Correlations {.tabset}

How to read: 

- Blue indicates positive correlation. Red indicates negative correlation.
- The darker the blue or red showcases the strength of the correlation. 
- Rule of thumb: Anything +/- 0.30 indicates the two metrics are correlated. Anything below this threshold should not be considered

```{r include=FALSE}
correlation_matrix_plot <- ggstatsplot::ggcorrmat(pch="",
  data = sample_data_tbl,
  type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
  colors = c("darkred", "white", "steelblue") # change default colors
)
```
```{r}
correlation_matrix_plot
```

&nbsp;

----

&nbsp;

## 2.0 Lag Correlations {.tabset}

How to read: 

- If point falls between the two dotted blue lines, no correlation.
- If point falls above or below the dotted blue lines, correlation.
- Hover over to see the correlation coefficient, if correlation is +/- 0.4 this indicates a mid/strong correlation.

Rule of thumb, Lag 1 (1 month) and 2 (2 month) would be good bets. Lag 6 if you think a 6-month lag is beneficial. And in rare cases Lag 12 if you think a 1-year lag is beneficial.

```{r}
sample_data_tbl %>%
  plot_acf_diagnostics(date, leads, 
                       .ccf_vars = c(spend_paid_search, spend_paid_social, spend_programmatic, spend_email),
                       .facet_ncol = 2,
                       .lags = 12)
```

&nbsp;

----

&nbsp;


## 3.0 Correlation Scatterplots {.tabset}

What to look out for:

- Linear relationships would be shown if all points are in a tight diagonal line. If increasing from left to right = positive linear relationship. If decreasing from the left to right = negative linear relationship
- If the points are all scattered, this is a result of inconclusive linear relationship. Not linear.

```{r include=FALSE}
scatterplot_correlated_tbl <- sample_data_tbl %>% 
  select (leads, spend_paid_search, spend_paid_social, spend_programmatic, spend_email)
correlation_scatter_plot <- scatterplot_correlated_tbl %>% 
  ggpairs()
```

```{r}
correlation_scatter_plot
```

&nbsp;

----

&nbsp;

## 4.0 Correlation Funnels

Purpose: What happens to the variables when it is high or low. For example when sales are high, is air travel also high? and bank loan rates are low?

### All In {.tabset}

```{r include=FALSE}
#load correlation script
root.dir = rprojroot::find_rstudio_root_file()
corr_script = paste(root.dir, "/00_scripts/plot_cor.R", sep='')
source(corr_script)

#create recipe for the discretize object
discretize_obj <- recipe(~ ., data = sample_data_tbl) %>%
  step_zv(all_predictors()) %>%
  step_discretize(all_numeric(), options = list(min_unique = 1)) %>%
  step_dummy(all_nominal(), one_hot = TRUE) %>%
  prep()

#bake
dist_tbl <- bake(discretize_obj, new_data = sample_data_tbl)

# write.csv(dist_tbl, file = "binned_distribution_table.csv", row.names = FALSE)

```

#### Sales High
```{r include=FALSE}
# Data manipulation.
correlation_bin4_tbl <- dist_tbl %>%
  select(-c (date)) %>%
  get_cor(leads_bin4, fct_reorder = T, fct_rev = T) %>%
  filter(abs(leads_bin4) >= 0.02) %>%
  mutate(
    relationship = case_when(
      leads_bin4 > 0 ~ "Supports",
      TRUE ~ "Contradicts"
    )
  ) %>%
  mutate(feature_text = as.character(feature)) %>%
  separate(feature_text, into = "feature_base", sep = "_", extra = "drop") %>%
  mutate(feature_base = as_factor(feature_base) %>% fct_rev())

# Visualization of discretized corr. 
correlation_bin4_plot <- correlation_bin4_tbl %>%
  ggplot(aes(leads_bin4, feature_base, color = fct_rev(relationship))) +
  geom_point() +
  geom_label(aes(label = feature), vjust = -0.5) + 
  expand_limits(x = c(-0.4, 0.4)) +
  theme_tq() +
  scale_color_tq() +
  labs(
    title = paste("Correlation Analysis: If Sales is HIGH"),
    subtitle = "How to read? Top correlated features are in the upper right, they then decrease in order downward (funnel)"
  )
```
```{r}
correlation_bin4_plot
```

&nbsp;

----

&nbsp;

#### Sales Low
```{r include=FALSE}
# Data manipulation.
correlation_bin1_tbl <- dist_tbl %>%
  select(-c (date)) %>%
  get_cor(leads_bin1, fct_reorder = T, fct_rev = T) %>%
  filter(abs(leads_bin1) >= 0.02) %>%
  mutate(
    relationship = case_when(
      leads_bin1 > 0 ~ "Supports",
      TRUE ~ "Contradicts"
    )
  ) %>%
  mutate(feature_text = as.character(feature)) %>%
  separate(feature_text, into = "feature_base", sep = "_", extra = "drop") %>%
  mutate(feature_base = as_factor(feature_base) %>% fct_rev())

# Visualization of discretized corr. 
correlation_bin1_plot <- correlation_bin1_tbl %>%
  ggplot(aes(leads_bin1, feature_base, color = fct_rev(relationship))) +
  geom_point() +
  geom_label(aes(label = feature), vjust = -0.5) + 
  expand_limits(x = c(-0.4, 0.4)) +
  theme_tq() +
  scale_color_tq() +
  labs(
    title = paste("Correlation Analysis: If Sales is HIGH"),
    subtitle = "How to read? Top correlated features are in the upper right, they then decrease in order downward (funnel)"
  )
```
```{r}
correlation_bin1_plot
```

&nbsp;

----

&nbsp;
